---
title: "Data Analysis Project"
author: "Cooper, Oliver, Michael"
date: "November, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 
In this notebook, we are analyzing Maryland Early Voting turnout data (https://elections.maryland.gov/press_room/index.html)

## Load libraries
Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor) #tidyverse, lubridate, and janitor are the only libraries we need for this project
```

## Load and Cleaning Data
In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

## Analysis 

The source of the data is releases and election reports from the Maryland State Board of Elections. There aren't any obvious limitations or flaws, other than the inability to answer questions involving other states with the current data set due to it just being Maryland data. 
```{r}
md_ev_gub_14 <- clean_names(read_csv("dataproj/md_ev_gub_14.csv")) #MD early voting for the gubernatorial election of 2014

md_ev_pres_16 <- clean_names(read_csv("dataproj/md_ev_pres_16.csv")) #MD early voting for the presidential election of 2016

md_ev_gub_18 <- clean_names(read_csv("dataproj/md_ev_gub_18.csv")) #MD early voting for the gubernatorial election of 2018

md_ev_pres_20 <- clean_names(read_csv("dataproj/md_ev_pres_20.csv")) #MD early voting for the presidential election of 2020

md_ev_gub_22 <- clean_names(read_csv("dataproj/md_ev_gub_22.csv")) #MD early voting for the gubernatorial election of 2022

md_ev_pres_24 <- clean_names(read_csv("dataproj/md_ev_pres_24.csv")) #MD early voting for the presidential election of 2024
```

## Sentences to Engineer
In this notebook, we are working to answer five questions about early voters in Maryland.

### Question 1
* **Question**: How have early voting numbers changed over the last ten years?

We want to identify general trends in the data, including changes and trends based on events is the news. This is a common thing that other journalists have done - checking trends - but we could differ from them by doing a timeline with events like political rhetoric or other topics.

* **Analysis summary**: Our analysis shows that in general, early voting has increased across counties in Maryland, with some like Washington and Calvert counties more than tripling their early voting turnout over the 10-year stretch. One thing that is a little surprising is the sharp decline in EV in Baltimore City, which began before the pandemic and seems to only now be slightly recovering. The only other county jurisdictions with declines from 2014-2024 are Somerset, Dorchester, Kent, and Anne Arundel, which do not have much in common with Baltimore City. The gubernatorial races tend to draw less early voting, which is not surprising because they draw less voting overall. Three counties did not have a single early voter in the 2022 race.

```{r}
md_ev_gub_14 <- clean_names(read_csv("dataproj/md_ev_gub_14.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_14=sum(total, na.rm=TRUE)
  )

md_ev_pres_16 <- clean_names(read_csv("dataproj/md_ev_pres_16.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_16=sum(total, na.rm=TRUE)
  )

md_ev_gub_18 <- clean_names(read_csv("dataproj/md_ev_gub_18.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_18=sum(total, na.rm=TRUE)
  )

md_ev_pres_20 <- clean_names(read_csv("dataproj/md_ev_pres_20.csv")) |>
  mutate(total=total_day1+total_day2+total_day3+total_day4+total_day5+total_day6+total_day7+total_day8) |>
  mutate(total=as.numeric(total)) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_20=sum(total, na.rm=TRUE)
  )

md_ev_gub_22 <- clean_names(read_csv("dataproj/md_ev_gub_22.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_22=sum(total, na.rm=TRUE)
  )
   
md_ev_pres_24 <- clean_names(read_csv("dataproj/md_ev_pres_24.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_24=sum(total, na.rm=TRUE)
  )

md_ev_combined <- md_ev_pres_24 |>
  left_join(md_ev_gub_22, join_by(county_name)) |>
  left_join(md_ev_pres_20, join_by(county_name)) |>
  left_join(md_ev_gub_18, join_by(county_name)) |>
  left_join(md_ev_pres_16, join_by(county_name)) |>
  left_join(md_ev_gub_14, join_by(county_name)) |>
  mutate(pct_change=100*(total_24-total_14)/total_14) |>
  arrange(desc(pct_change))

md_ev_combined

```

### Question 2
* **Question**: How is early voting data impacted by which party is projected to win the election (currently winning in the polls) at the time?

Early voting data is very important - looking at trends and differences in this data can tell you a lot about an election. However, we think it’s newsworthy to see how people’s confidence in their party affected early voting, if at all.

 **2024 - totaling only dems and reps - 110593, dems=63.7%, reps=36.3%**

**2020 - totaling only dems and reps - 143702, dems=58.5%, reps=41.5**

**2016 - totaling only dems and reps - 72730, dems=65.2%, reps=34.8%**

**Final polling numbers - 2016: Clinton 45.7, Trump 41.8 - 2020: Biden 51.8, Trump 43.4 - 2024: Harris 48, Trump 46.8**

**Cross-referencing these numbers with 538's final polling numbers, it seems that there's not much correlation between early voting data and the final polling numbers. If anything, the closer the final polls are, the larger the gap is between the two parties in registered voters. However, there is more of a correlation between the total number of early voters and the polls - the better the polls look for democrats, the higher the final number of early voters**

```{r}
md_ev_pres_24 |>
  group_by(party_code)|>
  summarise(count = n())
```

``` {r}
md_ev_pres_16 |>
  group_by(party_code)|>
  summarise(count = n())


```

``` {r}
md_ev_pres_20 |>
  group_by(party_code)|>
  summarise(count = n()) 



```

### Question 3
* **Question**: Were there any major shifts in terms of early voting during the pandemic as far as the 2020 presidential election?

The pandemic had a tremendous impact on life for just about everyone, but also an incredible impact on election coverage, conversation, and activism. Understanding if there was a difference in this election year compared to other years is very important for that exact reason. There has been reporting on this effect on a national scale. In terms of Maryland, it’s now just used as a comparison to the current election race.

* **Analysis summary**: While total early voting certainly spiked for the 2020 election, our analysis finds that it did not increase in every county. Prince George's County, for example, had higher early voting in both 2016 and 2024. Other counties of varying locations and population sizes experienced different trends in their early voting when comparing 2020 to '16 and '24, painting a picture of almost random output.

```{r}
md_ev_pres_16 <- clean_names(read_csv("dataproj/md_ev_pres_16.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_16=sum(total, na.rm=TRUE)
  )

md_ev_pres_20 <- clean_names(read_csv("dataproj/md_ev_pres_20.csv")) |>
  mutate(total=total_day1+total_day2+total_day3+total_day4+total_day5+total_day6+total_day7+total_day8) |>
  mutate(total=as.numeric(total)) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_20=sum(total, na.rm=TRUE)
  )
    
md_ev_pres_24 <- clean_names(read_csv("dataproj/md_ev_pres_24.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(county_name, total) |>
  group_by(county_name) |>
  summarise(
    total_24=sum(total, na.rm=TRUE)
  )

md_ev_pres_combined <- md_ev_pres_24 |>
  left_join(md_ev_pres_20, join_by(county_name)) |>
  left_join(md_ev_pres_16, join_by(county_name)) |>
  arrange(desc(total_20))

ggplot(md_ev_pres_combined, aes(x=total_24, y=total_20, color=total_16)) +
  geom_point()

```

### Question 4
* **Question**: How does the early voting turnout in Maryland compare to other states of similar population size and geographic location?

This specifically addresses how local states are doing with voter turnout (essentially the DMV and just outside of it). Having this information will show interesting trends and also any discrepancies of voters that are geographically close to Maryland. There has been extensive research and journalistic work done on states early voter turnout but not so much based on geographic location and population size.

* **Analysis summary**: Early voting data was pervasive in Pennsylvania and Delaware, so we looked at Virginia as our neighborly comparison. As we can see from our analysis, a higher percentage of VA residents participated in early voting in 2024. Since Virginia's population is also higher than Maryland's, this means that almost 150,000 more people early voted in VA than in MD.

```{r}
md_ev_pres_24 <- clean_names(read_csv("dataproj/md_ev_pres_24.csv")) |>
  mutate(total=day1+day2+day3+day4+day5+day6+day7+day8) |>
  select(total) |>
  summarise(
    total_md=sum(total, na.rm=TRUE)
  ) #finding total EV for 2024, same code as Q1 and Q3

va_ev_pres_24 <- clean_names(read_csv("dataproj/va_ev_pres_24.csv")) |>
  select(early_voting) |>
  summarise(
    total_va=sum(early_voting, na.rm=TRUE)
  ) #finding total EV for Virginia in 2024, grabbed csv file from their elections website just like in MD. PA and DE had data for registration and results by candidate, but didn;t include columns for early voting in their data. We could only find absentee ballots, provisional ballots, and in-person votes.

md_va_24 <- md_ev_pres_24 |>
  cross_join(va_ev_pres_24) #cross join since they do not share a column name

library(tidyr)
library(dplyr) #both needed for cleaning the combined dataframe so that it can be suitable for ggplot
md_va_24 <- md_va_24 |>
  mutate(population_md=6180253) |> #as of July, 2023 via https://www.census.gov/quickfacts/fact/table/MD/PST045223
  mutate(population_va=8715698) |> #as of July, 2023 via https://www.census.gov/quickfacts/fact/table/VA/PST045223
  mutate(pct_md=total_md/population_md*100) |>
  mutate(pct_va=total_va/population_va*100) |> #new columns showing how much of the population did early voting
  pivot_longer(cols=everything(), names_to="key", values_to="value") |> #pivot_longer() function initiates the process of cleaning/flipping the dataframe so that it becomes sutiable for ggplot. "cols=everything()" makes sure we are working with all of our columns. "names_to='key'" essentially creates a bank of the column names that we will use next. same for the numbers in "value".
  separate(key, into=c("metric", "state"), sep="_") |> #separate() function allows us to clean our column names, which we know are now stored in "key". We take everything before the underscore and call it a metric (total, population, and pct) and make everything after the underscore a state (md, va). this would not work if our column names did not all follow the structure of metric_state.
  pivot_wider(names_from=metric, values_from=value) #now we simply pivot the dataframe to where the column names are the ones we split from their underscores in "metric", the rows are the other half of those underscored titles, and the numbers we stored in "value" beforehand return.

md_va_24

md_va_24 |>
  ggplot() +
  geom_bar(aes(x=state, weight=total, fill=pct)) #the color of the bars reflects the percentage of the state's population that early voted.
```


### Question 5
* **Question**: How is early voting different based on the state's traditional political support? 

Understanding how political party affiliation is involved in early voting data will give us an insight on early voting trends based on political support in different states.

** Maryland got a close percentage of democratic support in the presidential race as South Dakota had republican support, so I'll be comparing the two. Looking at the party affiliation of the early voters in both states, you can see that South Dakota's early vote was much more one-sided than Maryland's, with republicans outnumbering democrats more than 3-1 in early voting, compared to Maryland, which was closer to 2-1. I think this tells a story that lines up with the final results - as indicated by the early voting distribution in two states that had close to the same percentage of the total vote for opposite candidates.  **

```{r}
md_early_voting_support <- clean_names(read_csv("md2024evbyparty.csv"))

md_early_voting_support |>
    ggplot() +
  geom_bar(aes(x=x1, weight=total_votes, fill=total_votes))+labs(x = "Party Affiliation")

sd_early_voting_support <- clean_names(read_csv("sd2024evbyparty.csv"))

sd_early_voting_support |>
      ggplot() +
  geom_bar(aes(x=x1, weight=total_votes, fill=total_votes))+labs(x = "Party Affiliation")
  
```


-30-