---
title: "Reverse Engineering Project"
author: "Michael, Dylan, Oliver"
date: "Oct. 5"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, "Why Many Americans don’t Vote"(https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor) 
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

The source of the data was 538.com, which used a voter file combined with survey responses to classify voters and their answers to questions. Didn't really come across many issues, but all of the data is numeric values. 

```{r}
Story_data <- read_csv("~/Downloads/nonvoters_data.csv")

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: .“Every election, millions of Americans go through a similar thought process and, it turns out, lots of people feel like Brown: They think voting doesn’t matter or isn’t worth their time.”
* **Analysis summary**:  611 people fit that category out of an original dataset of 5,836 - roughly 10 percent, which would equate to millions of people with the correlation to the US voting population. This demonstrates that if the dataset represents all americans, millions think that voting doesn't matter or isn't worth their time. 

```{r}
filtered_data <- Story_data
filtered_data |>
  filter(Q26 >= 2) |>
  filter(Q29_2 == 1 | Q29_3 == 1 | Q29_4 == 1)
print(filtered_data)

```

### Sentence 2

* **Sentence text**: “In the survey, we asked voters who have missed at least one national election — which included some people who almost always vote — why they didn’t cast a ballot. Nearly a quarter cited some of the structural barriers we mentioned above. But another 31 percent said that they decided not to vote because they disliked the candidates or they thought nothing would change as a result of the election (26 percent).”


* **Analysis summary**: Out of 1,342 people who say they sometimes, rarely, or don't vote in national elections, 765 say that they either dislike the candidates or that nothing will change for them. 403 say they dislike the candidates, while 362 say nothing will change as a result. This is sinilar to the original numbers presented in the text, as 403 is 30% of 1342 and 362 is 27%.

```{r}
filtered_data <- Story_data
filtered_data |>
  filter(Q26 >= 2) |>
  filter(Q29_1 == 1 | Q29_3 == 1)|>
print(filtered_data)


```

### Sentence 3

* **Sentence text**: [A substantial chunk of respondents said they don’t think either party wants people like them to vote. (Twenty-three percent said this of Democrats and 31 percent of Republicans.) Black and Hispanic people were especially likely to say the Republicans don’t want people like them to vote: Fifty-four percent of Black respondents and 35 percent of Hispanic respondents agreed with this statement, compared to just 26 percent of white respondents.]
* **Analysis summary**: The result of the analysis confirm the findings pretty well. Q14, which represents the republican party, shows that 33% of respondents say they don't think republicans want them to vote, similar to the 31% in the article. Q15, which is the democratic party, got a number of 21%, but accounting for the people that didn't vote the number should be similar to the 23% in the article. 

```{r}
Story_data |>
  group_by(Q14) |>
summarise(count = n()) |>
mutate(percent = (count/sum(count)*100))
```
``` {r}
Story_data |>
  group_by(Q15) |>
summarise(count = n()) |>
mutate(percent = (count/sum(count)*100))
```
### Sentence 4

* **Sentence text**: “They were also likelier than those who vote more regularly to say they couldn’t get off work to vote...
Also because he doesn't trust mail-in ballots. "The last time I tried to vote by mail, it didn't get counted-it was too late," he said. As a result, Sanchez told us, "I don't know where my ballot is going to end up, if I don't see it put into the voting machine."
* **Analysis summary**: When filtering the data, 229 respondents chose 5 on the survey, which has the answer of they wanted to vote but had work/other obligations or forgot. Furthermore, in question 19 asks respondents what they think would make it easier for them to vote. The responses that have to do with automatically being registered to vote and recieving a mail-in ballot which had thousands of respondents say that this would make voting a lot easier for them. Comparing these different answers shows what would make voting easier for those who don't know about mail-in deadlines. 

```{r}
filtered_data_4 <- Story_data |>
  select(Q29_5)|>
  filter(Q29_5 == 1)
filtered_data_4
  

```

### Sentence 5

* **Sentence text**: [ Black and Hispanic voters are also more likely to experience hurdles, perhaps in part because there tend to be fewer polling places in their neighborhoods. About 24 percent of Black respondents said that they had to stand in line for more than an hour while voting, and Hispanic respondents were more likely to say they had trouble accessing the polling place or couldn’t get off work in time to vote.”]
* **Analysis summary**: 248/932, or 26% of black respondents, said that they had to stand in line for more than an hour while voting, which is similar to the 24% number in the article. 130/813 Hispanics said they had trouble accessing the polling place or couldn't get off work in time to vote. 

```{r}
# Put code to reverse engineer sentence here
filtered_data <- Story_data
filtered_data |>
  filter(race == 'Black') |>
  filter(Q18_8 == 1)|>
print(filtered_data)

filtered_data <- Story_data
filtered_data |>
  filter(race == 'Hispanic') |>
  filter(Q18_2 == 1 | Q18_7 ==1)|>
print(filtered_data)

```

-30-