---
title: "Reverse Engineering Project"
author: "Michael, Dylan, Oliver"
date: "Oct. 5"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, "Why Many Americans don’t Vote"(https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor) 
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

The source of the data was 538.com, which used a voter file combined with survey responses to classify voters and their answers to questions. Didn't really come across many issues, but all of the data is numeric values. 

```{r}
Story_data <- read_csv("~/Downloads/nonvoters_data.csv")

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: .“Every election, millions of Americans go through a similar thought process and, it turns out, lots of people feel like Brown: They think voting doesn’t matter or isn’t worth their time.”
* **Analysis summary**:  611 people fit that category out of an original dataset of 5,836 - roughly 10 percent, which would equate to millions of people with the correlation to the US voting population. This demonstrates that if the dataset represents all americans, millions think that voting doesn't matter or isn't worth their time. 

```{r}
filtered_data <- Story_data[Story_data$Q26 %in% c(2, 3, 4), ]
view(filtered_data)

filtered_data <- filtered_data[
  filtered_data$Q29_2 == 1 | 
    filtered_data$Q29_3 == 1 | 
    filtered_data$Q29_4 == 1, 
]
print(filtered_data)

```

### Sentence 2

* **Sentence text**: “In the survey, we asked voters who have missed at least one national election — which included some people who almost always vote — why they didn’t cast a ballot. Nearly a quarter cited some of the structural barriers we mentioned above. But another 31 percent said that they decided not to vote because they disliked the candidates or they thought nothing would change as a result of the election (26 percent).”


* **Analysis summary**: Out of 1,342 people who say they sometimes, rarely, or don't vote in national elections, 765 say that they either dislike the candidates or that nothing will change for them. 403 say they dislike the candidates, while 362 say nothing will change as a result. This is sinilar to the original numbers presented in the text, as 403 is 30% of 1342 and 362 is 27%.

```{r}
filtered_data <- Story_data[Story_data$Q26 %in% c(2, 3, 4), ]
view(filtered_data)

filtered_data <- filtered_data[
  filtered_data$Q29_1 == 1 |
    filtered_data$Q29_3 == 1,
]
print(filtered_data)

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
filtered_data <- Story_data[Story_data$Q14 %in% c(4,5), ]
view(filtered_data)

filtered_data <- filtered_data[
  filtered_data$Q29_1 == 1 |
    filtered_data$Q29_3 == 1,
]
print(filtered_data)

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-